\subsection{XGBoost}
\subsubsection{Definition}

XGBoost (Extreme Gradient Boosting) is a powerful machine learning algorithm that builds an ensemble of decision trees using the gradient boosting framework. It is widely recognized for its speed, scalability, and high accuracy in handling structured data for both classification and regression problems.

One of the models that attracted the most attention when choosing which ones would be used to solve the toxicity classification task was XGBoostClassifier. According to [], XGBoost stands for Extreme Gradient Boosting and is based on the use of multiple decision and regression trees to create an ensemble method with greater predictive capacity. In this model, scores are assigned to the examples based on the leaf in which they are classified by each tree, and then those scores are added together. This approach is similar to that used in Random Forests, except that in this case, training is done using Boosting. The objective of the method is to learn both the structure and the scores of the trees. In each training iteration, the goal is to find the tree that contributes most to optimizing the objective, which can vary depending on the task. It can be said that the trees in the ensemble method complement each other to solve the task for which they are trained.

In [], the ensemble methods tested were AdaBoost and Random Forest, with which an accuracy of 95.518 was achieved using Random Forest. On the other hand, in [], using this same classifier, an accuracy of 91.88 was achieved. Finally, in [], also with this classifier, an accuracy of 85.1 was obtained. From these results, these scores can be taken as a reference and the search for some improvement can be established.

